\Large \bf Cyclone: Warp Speed Replication for Key Value Stores

### 2017-04-21 paper/paper.tex
Updates to paper.

\begin{abstract}
Persistent key value stores are rapidly becoming an important component of many
distributed data serving solutions with innovations targeted at taking advantage
of growing flash speeds. Unfortunately their performance is hampered by the need
to maintain and replicate a write ahead log to guarantee availability in the
face of machine and storage failures. Cyclone offers a solution to this problem
by leveraging a small amount of directly attached non-volatile memory (NVM) such
as NVDIMMs to maintain a two level log: the upper level in a small amount of NVM
draining into a larger lower level log maintained on flash. Replicating the
smaller log kept in NVMs transforms the replication problem into a packet
switching problem as the leader sends the exact same packet to all follower
replicas and we solve the problem as such by using the RAFT protocol implemented
as a software multicast module on the Data Plane Development Kit (DPDK). Second,
we leverage the observation that key value store interfaces are commutative for
operations to different keys to scale our two level log horizontally into a
number of physical two level logs - with a hash of the key being used to choose
the physical log. Each log is replicated by an independent instance of RAFT,
thereby leveraging the scalability afforded by classic software packet switching
when using multiple CPU cores. Cyclone is able to replicate millions of small
updates a second using only commodity 10 gigabit ethernet adapters and improves
the performance and availability of RocksDB, a popular persistent key value
store by an order of magnitude when compared to its own write ahead log without
replication.
\end{abstract}

@@

These key value stores however tend to ignore an important component: the write
ahead log. A machine or storage failure leading to total loss or temporary
unavailability of data is unacceptable in many internet scale services where
high availabilty and revenue are often interconnected. Key value stores
therefore usually incorporate support for a write ahead log that if replicated
and kept durable for every appended update provides the necessary high
availability. Unfortunately the write ahead is a performance achilles heel for
these systems - eclipsing much of the work on improving performance of the LSM
component. To illustrate the impact of the write ahead log, consider
Figure~\ref{fig:problem}. The line marked Rocksdb shows the performance of
Rocksdb without the write ahead log. The performance when persisting the write
ahead log without replicating it is shown as the line marked 'rocksdb/WAL'. The
line marked 'rocksdb/3 way rep.' is for simply replicating the log three ways
without persisting it, using RAFT~\cite{raft} running over TCP/IP. Either
persisting every update to the log or replicating it using TCP/IP causes
performance to drop by an order of magnitude (observe the log scale on the
x-axis). It is therefore no suprise that deployments of Rocksdb often turn off
the write ahead log~\cite{samza}. On the other side of the spectrum, key value
store research such as FloDB~\cite{flodb} turns off the write ahead log to be
able to showcase benefits of sophisticated extensions to LSM datastructures.

\begin{figure}
\centering
\includegraphics[scale=0.6]{results2/problem.pdf}
\caption{Rocksdb Write Ahead Logging Impact}
\label{fig:problem}
\end{figure}

@@

Cyclone makes use of a limited amount of attached NVM using two ideas. First, it
leverages the observation that with persistence being a first class property of
a subset of directly addressable DRAM, replicating the same log entry from the
leader to follower replicas is no different from multicast using software packet
switching techniques. It directly leverages this observation to implement a
version of RAFT that replicates the NVM log using classic packet switching
techniques such as separation of control and data planes. Second, Cyclone uses a
two level log structure where it drains entries from the NVM log in IO friendly
batches to a log placed on a flash drive. This allows Cyclone to use only a
limited amount of NVM reasoning that current forms of NVM such as NVDIMMs are
more expensive than DRAM. Our experiments use an NVM footprint of a mere 64 MB
to achieve performance comparable to placing the entire log in NVM. This allows
Cyclone to be used in deployments where the key value store is maintained
entirely in main memory rather than as a multi-level LSM tree structure, with
the log serving as the source for recovery in the event of failure.

@@

\section{Two Level Log}
\label{sec:2level}
A fundamental challenge in building an efficient replicated log is dealing with
network overheads. Cyclone's construction is predicated on the observation that
in the common case, protocols such as RAFT simply multicast the same entry
received from a client, from the leader replica to all follower replicas. This
suggests that an efficient way to implement a replicated log is to treat it as a
software packet switching problem. However peristing logs on flash represents an
impediment in that block storage devices are not good candidates to hold packets
being software switched. Our solution is to split the log into two levels. The
upper level is held in persistent directly attached non-volatile memory (NVM)
such as NVDIMMs. The lower level is held on a flash SSD - we term the lower
level log as the flashlog. We periodically drain entries in batches from the NVM
log to the backing flashlog. An important consequence of this two level approach
is that we need only a limited amounts of NVM with the bulk of the log residing
on flash - an arrangement that maps well to the current relative costs of the
two types of memory. We begin by describing the NVM log and how we implement the
RAFT consensus protocol in the packet switching paradigm around it. We then
describe the flashlog and how we drain entries in batches from the NVM level to
the flash level.

@@

\begin{figure}
\centering
\includegraphics[scale=0.3]{figures2/nvm_log.pdf}
\caption{NVM Log Structure}
\label{fig:nvm_log}
\end{figure}

@@

\begin{figure}
\begin{verbatim}
    event_receive_client_req()
    {
      if(!check_is_leader()) {
        drop_msg
        return
      }
      Prepend raft log term and index
      Persist to local log
      Transmit to follower replicas
    }
\end{verbatim}
\caption{Cyclone Control Plane Fragment}
\label{fig:control_plane}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[scale=0.3]{figures2/network_packet.pdf}
  \caption{Cyclone Network Packet Layout}
  \label{fig:packet_layout}
\end{figure}

We note that by discrding stream oriented protocols such as TCP, the network
layer is not longer reliable or ordered. This is not a problem for RAFT that is
designed to tolerate network failures resulting in drops or reordering of
network packets. However, this rules out consensus protocols such as Zookeeper
Atomic Broadcast (ZAB~\cite{zab}) that depend on a reliable stream-oriented
network layer to function correctly.

Finally, we turn our attention to the persistence step in
Figure~\ref{fig:control_plane}. RAFT requires that the log entry be persisted
before it is multicast out to follower replicas. Since the NVM is directly
attached we do this by executing a cacheline flush ({\tt clflush}) instruction
for every cacheline in the packet and the pointer in the pointer buffer to
persist these via the memory bus. This is the one exception to our otherwise
clean separation between the control and data plane - necessitated by the fact
that at the moment NICs do not support persistent memory. This is not too
onerous a burden because we can use the newly introduced {\tt
  clflush-opt}~\cite{clflush_opt} instruction specifically intended to
efficiently flush to persistent memory without the overhead of the serialization
normally introduced by {\tt clflush}. This allows one to hit full memory
bandwidth on present generation platforms, a quantity in excess of 200 Gb/s per
core, well above the near term speeds of network interface cards.

@@

In order to recover from a crash, we make two important assumptions about the
underlying SSD. First, we assume that 4KB is the \emph{minimum} atomic unit for
updating pages on the SSD even under power failure i.e. there are no shorn
writes on a 4KB page~\cite{shorn_writes}. We also assume that the SSD has power
loss data protection meaning that writes cached in the drive's volatile cache
are written to the SSD using a backup capacitor in the event of power failure -
a property of many data center class SSDs today, including the Intel DC P3600
SSD~\cite{ssd_spec} we use in our evaluation. Together these two assumptions
mean that we can recover a consistent prefix of the log on a power failure.

@@

The flow for an update operation in Cyclone is as follows. The network packet is
received by the control plane for the targeted physical log. The packet is then
immediately sent out for replication as described in the previous section. When
responses from a majority quorum is received, the request is passed to the
targeted KV store core for execution. The KV store core process the request and
sends back the response to the client. We note that replication is pipelined and
decoupled from execution and thus multiple updates are usually in
flight on the network for replication at the same time.

@@

We note that commutativity as assumed by Cyclone is an \emph{API level
  property}. There are no assumption as to how application cores synchronize to
access the shared key-value state or the underlying data structures for holding
key-value pairs. We emphasize that this means replicas ultimately have the same
content but not necessarily the same memory state.

@@

Our solution instead is to use a technique we termed ganged operations. It is
based on the observation that in the failure free case, executing a batched
write across multiple control plane cores and KV store cores does not look very
much different from a synchronous shared memory barrier. The challenge is to
effectively manage failures in replication without causing participating cores
to stop making progress. The first step is to reconfigure RAFTs leader election
protocol to ensure that the leaders for all the physical logs are co-located on
the same host. Clients always dispatch batched writes to a fixed control plane
core (called the injection core) which is then responsible for forwarding the
request to the participating control plane cores (for corresponding physical
logs). The injection core uses DPDKs packet cloning primitives to avoid
dataplane work while forwarding copies of the same batched write packet to
multiple control plane cores. The injection core also adds a ``nonce'' - a
unique timestamp to the packet. In addition, the injection core adds a unique
view number to the packet containing the term numbers of all the participating
RAFT instances (read from shared memory of the co-located leaders).  We describe
the remaining flow for ganged operations in terms of the pseudocode in
Figure~\ref{fig:ganged_ops}. We defer discussion of how we generate the nonce to
later in this section. We also assume that a unique barrier is allocated in
shared memory for each ganged operation. We discuss later how this is done
without using dynamic allocation.

In the absence of failures Figure~\ref{fig:ganged_ops} is straightforward. Once
replication is complete on all participating physical logs, the corresponding KV
store cores execute \\{\tt event\_replicated\_ganged\_op}. They use a bitmask for
the barrier setting the appropriate bit for themselves. Once all KV store cores
have arrived at the barrier the operation is executed. A distinguished leader
core (we use the minumum numbered participating KV store core) is
responsible for sending back the response to the client. Two types of failure
need to be considered. First, replication on the physical log corresponding to
one of the non-leader cores could fail. The
remaning cores would never progress past the barrier. To solve this the leader
core montors the remaning cores for their current view - which is the RAFT term
of the last executed log entry. If the view has moved past the view in the
packet, replication failed on that physical log. The leader then sets the bit
for the core that will no longer participate in the barrier to aid progress (a
technique we borrowed from lock-free data structures~\cite{lock-free}). It also
sets a failed flag to signal failure to the other participating cores. The
second failure case is that the leader core might not receive the operation if
it fails during replication from corresponding physical log. This case is
detected in a similar manner by the remaining cores by monitoring the published
view from the leader core in shared memory. We add that on a leadership change
due to timeout we enqueue and execute a nop on the physical log and all
associated KV store cores to ensure previous entries are committed - a corner
case in the RAFT protocol.

@@

We now turn our attention from the network component of Cyclone to the storage
one by adding the second level flashlog. We evaluate the impact of adding
block storage to our hitherto pure packet switching scenario in
Figure~\ref{fig:flashlog}. Batching entries from the top level NVDIMM log to the
second level flashlog is clearly beneficial as adding flash storage at the
second level has almost no impact on peak performance in Cyclone for small
entries. The situation however changes for larger
entries. Figure~\ref{fig:flashlog_512} shows that using a 512 byte payload has a
significant impact on peak throughput - it drops to approximately 350K
ops/sec. This corresponds to around 50K 4KB IOPS to the SSD to write out the
flashlog pages. The peak for the drive is 160K IOPS using a queue depth
(concurrency) of 128. With 32 application threads we expect a lower peak
throughput around 40K IOPS explaining our bottleneck at 50K IOPS.  It is
possible to tune our observed performance further by aligning the flush boundary
to increase the number of outstanding requests - we do not do so in this paper,
keeping Cyclone agnostic to the exact size of entry being replicated. A final
point about Figure~\ref{fig:flashlog_512} is that once we are past the storage
bottleneck the latency spike is dramatic and large enough to trigger Cyclone's
failure detector and repeated retries from the clients. There are - therefore -
no points on the ``knee'' of the curve as in the pure packet switched one level
log case.

@@

The final dimension we evaluate is of ganged operations. The primary purpose of
ganged operations is to avoid the need for distributed transactions to
manipulate what is a single shared memory image and therefore we were most
concerned about unloaded latency given the complexity of synchronizing different
replication quorums as well executing our rendezvous protocol on multiple
cores. We therefore setup an experiment where a single client - reflecting the
unloaded case - made ganged requests to the replicas. We varied the number of
cores participating in the request from one to the full complement of 32
application cores. Figure~\ref{fig:ganged} shows the results both using a single
level log as well as a two level log. The primary takeaway is that unloaded
latency increases slowly as we increase the number of active cores - to around
40 us from the baseline of 20 us. There are two causes of this. First, there is
a quicker rate of increase from 1 to 8 active cores, the reason being that this
corresponds to an increase in the number of physical logs that must be
synchronized on. There is some amount of synchronization involved in the
userspace DPDK driver for access to common resources for accessing the NIC and
this affects the ability to simultaneously issues replication messages on all
quorums. The smaller contribution to increasing latency that continues past the
8 core case is due to cacheline pingponging when executing the rendevous of
Figure~\ref{fig:ganged_ops}. Both these sources of latency drift could be
corrected using replication quorums mapped to dedicated NICs and more scalable
rendezvous designs (such as with machine aware broadcast
trees~\cite{broadcast_tree}). However we deemed the complexity of such
optimizations unnecessary. The added latency for most cases is well under the
extra round trip delay (including replication for all phases of a 2PC), which
would the minimum needed for a solution using distributed transactions via two
phase commit.

@@

Cyclone takes a software only approach to improving replication performance
using commodity network hardware. At the other end of the spectrum, Consensus in
a box~\cite{consensus_box} implements the Zookeeper atomic broadcast protocol
together with a NIC on an FPGA. That work ignored durability focusing purely on
replication performance. Latency numbers are impressive from the FPGA that cuts
out the path from CPU caches to the NIC and are reported to be as low as 3us for
a round trip from leader to follower replica compared to the 7us we observe with
DPDK. Notably however they reported 7us when using TCP/IP from the FPGA rather
than their own connection oriented network protocol. On the other hand peak
replication throughput appears to be 6M replications/sec on 66 Gbps of aggregate
network bandwith, comparable to the mark obtained by Cyclone on an aggregate 40
Gbps of network bandwidth. We believe the general applicability afforded by our
software only packet switching solution makes it a compelling alternative to an
FPGA based solution even given the larger latency.

@@

Cyclone provides a client library that automatically times out and retries
requests to deal with failure. Protection against repeated execution of the same
request if currently left to the key value store and client, since we felt that
exactly once semantics is optional when updates such as put operations are
idempotent. We have experimented with exactly once semantics as a basis for
distributed transactions across shards running Cyclone integrated key value
stores, along the lines of similar work~\cite{raft_lin}, using a fixed client
space and maintaining information about last seen sequence numbers from clients
in NVM. Distributed transactions on Cyclone shards however are outside the scope
of this paper.

\section{Conclusion}
Persistent key values stores struggle today to provide high availability using a
write ahead log without a significant impact to performance. Cyclone solves this
problem by leveraging a small amount of directly attached non-volatile memory
that transforms the replication problem into a software packet switching
problem. Cyclone and its Rocksdb bindings are slated for release to the open
source community and we hope both the code and the ideas in the system itself
will help key value store builders and researchers deal with their logging
problem.
\newcommand\myurl[2]{\url{#1}}
\bibliographystyle{plain}
\bibliography{paper}

>>>>

\begin{abstract}
Persistent key value stores are rapidly becoming an important component of many
distributed data serving solutions with innovations targeted at taking advantage
of growing flash speeds. Unfortunately their performance is hampered by the need
to maintain and replicate a write ahead log to guarantee availability in the
face of machine and storage failures. Cyclone offers a solution to this problem
by leveraging a small amount of directly attached non-volatile memory (NVM) such
as NVDIMMs to maintain a two level log: the upper level in a small amount of NVM
draining into a larger lower level log maintained on flash. Replicating the
smaller log kept in NVMs transforms the replication problem into a packet
switching problem as the leader sends the exact same packet to all follower
replicas. We solve this problem by using the RAFT protocol implemented
as a software multicast module on the Data Plane Development Kit (DPDK). In addition,
we leverage the observation that key value store interfaces are commutative for
operations to different keys to scale our two level log horizontally into a
number of physical two level logs. A hash of the key is used to choose
the physical log. Each log is replicated by an independent instance of RAFT,
thereby leveraging the scalability afforded by classic software packet switching
when using multiple CPU cores. Cyclone is able to replicate millions of small
updates a second using only commodity 10 gigabit ethernet adapters, improving
the performance and availability of RocksDB, a popular persistent key value
store by an order of magnitude when compared to its own write ahead log without
replication.
\end{abstract}

@@

These key value stores however tend to ignore an important component: the write
ahead log. A machine or storage failure leading to total loss or temporary
unavailability of data is unacceptable in many internet scale services where
high availabilty and revenue are often interconnected. Key value stores
therefore usually incorporate support for a write ahead log that if replicated
and kept durable for every appended update provides the necessary high
availability. Unfortunately the write ahead is a performance achilles heel for
these systems - eclipsing much of the work on improving the performance of the LSM
component. To illustrate the impact of the write ahead log, consider
Figure~\ref{fig:problem}. The line marked 'Rocksdb' shows the performance of
Rocksdb without the write ahead log. The performance when persisting the write
ahead log without replicating it is shown as the line marked 'rocksdb/WAL'. The
line marked 'rocksdb/3 way rep.' is for simply replicating the log three ways
without persisting it, using RAFT~\cite{raft} running over TCP/IP. Either
persisting every update to the log or replicating it using TCP/IP causes
performance to drop by an order of magnitude (note the log scale on the
x-axis). It is therefore no suprise that deployments of Rocksdb often turn off
the write ahead log~\cite{samza}. On the other side of the spectrum, key value
store research such as FloDB~\cite{flodb} turns off the write ahead log to be
able to showcase benefits of sophisticated extensions to LSM datastructures.

\begin{figure}
\centering
\includegraphics[scale=0.6]{results2/problem.pdf}
\caption{Rocksdb write ahead logging impact}
\label{fig:problem}
\end{figure}

@@

Cyclone makes use of a limited amount of attached NVM using two ideas. First, it
leverages the observation that with persistence being a first class property of
a subset of directly addressable DRAM, replicating the same log entry from the
leader to follower replicas is no different from multicast using software packet
switching techniques. It directly leverages this observation to implement a
version of RAFT that replicates the NVM log using classic packet switching
techniques such as separation of control and data planes. Second, Cyclone uses a
two level log structure where it drains entries from the NVM log in IO friendly
batches to a log placed on a flash drive. This allows Cyclone to use only a
limited amount of NVM following the reasoning that current forms of NVM such as
NVDIMMs are more expensive than DRAM. Our experiments use an NVM footprint of a
mere 64 MB to achieve performance comparable to placing the entire log in
NVM. This allows Cyclone to be used in deployments where the key value store is
maintained entirely in main memory rather than as a multi-level LSM tree
structure, with the log serving as the source for recovery in the event of
failure.

@@

\section{Two Level Log}
\label{sec:2level}
A fundamental challenge in building an efficient replicated log is dealing with
network overheads. Cyclone's construction is predicated on the observation that
in the common case, protocols such as RAFT simply multicast the same entry
received from a client, from the leader replica to all follower replicas. This
suggests that an efficient way to implement a replicated log is to treat it as a
software packet switching problem. However peristing logs on flash represents an
impediment in that block storage devices are not good candidates to hold packets
being software switched. Our solution is to split the log into two levels. The
upper level is held in directly attached non-volatile memory (NVM)
such as NVDIMMs. The lower level is held on a flash SSD - we term the lower
level log as the flashlog. We periodically drain entries in batches from the NVM
log to the backing flashlog. An important consequence of this two level approach
is that we need only a limited amounts of NVM with the bulk of the log residing
on flash - an arrangement that maps well to the current relative costs of the
two types of memory. We begin by describing the NVM log and how we implement the
RAFT consensus protocol in the packet switching paradigm around it. We then
describe the flashlog and how we drain entries in batches from the NVM level to
the flash level.

@@

\begin{figure}
\centering
\includegraphics[scale=0.3]{figures2/nvm_log.pdf}
\caption{NVM log structure}
\label{fig:nvm_log}
\end{figure}

@@

\begin{figure}
\begin{verbatim}
    event_receive_client_req()
    {
      if(!check_is_leader()) {
        drop_msg
        return
      }
      Prepend raft log term and index
      Persist to local log
      Transmit to follower replicas
    }
\end{verbatim}
\caption{Cyclone control plane fragment}
\label{fig:control_plane}
\end{figure}

\begin{figure}
  \centering
  \includegraphics[scale=0.3]{figures2/network_packet.pdf}
  \caption{Cyclone network packet layout}
  \label{fig:packet_layout}
\end{figure}

We note that by discarding stream oriented protocols such as TCP, the network
layer is no longer reliable or ordered. This is not a problem for RAFT that is
designed to tolerate network failures resulting in drops or reordering of
network packets. However, this rules out consensus protocols such as Zookeeper
Atomic Broadcast (ZAB~\cite{zab}) that depend on a reliable stream-oriented
network layer to function correctly.

Finally, we turn our attention to the persistence step in
Figure~\ref{fig:control_plane}. RAFT requires that the log entry be persisted
before it is multicast out to follower replicas. Since the NVM is directly
attached we do this by executing a cacheline flush ({\tt clflush}) instruction
for every cacheline in the packet and the pointer in the pointer buffer to
persist these via the memory bus. This is the one exception to our otherwise
clean separation between the control and data plane - necessitated by the fact
that at the moment NICs do not support persistent memory. This is not too
onerous a burden because we can use the newly introduced {\tt
  clflush-opt}~\cite{clflush_opt} instruction specifically intended to
efficiently flush to persistent memory without the overhead of the serialization
normally introduced by {\tt clflush}. This allows us to hit full memory
bandwidth on present generation platforms, a quantity in excess of 200 Gb/s per
core, well above the near term speeds of network interface cards.

@@

In order to recover from a crash, we make two important assumptions about the
underlying SSD. First, we assume that 4KB is the \emph{minimum} atomic unit for
updating pages on the SSD even under power failure i.e. there are no shorn
writes (otherwise known as torn writes) on a 4KB page~\cite{shorn_writes}. We
also assume that the SSD has power loss data protection meaning that writes
cached in the drive's volatile cache are written to the SSD using a backup
capacitor in the event of power failure - a property of many data center class
SSDs today, including the Intel DC P3600 SSD~\cite{ssd_spec} we use in our
evaluation. Together these two assumptions mean that we can recover a consistent
prefix of the log on a power failure.

@@

The flow for an update operation in Cyclone is as follows. The network packet is
received by the control plane for the targeted physical log. The packet is then
immediately sent out for replication as described in the previous section. When
responses from a majority quorum is received, the request is passed to the
targeted KV store core for execution. The KV store core processes the request
and sends back the response to the client. We note that replication is pipelined
and decoupled from execution and thus multiple updates are usually in flight on
the network for replication at the same time.

@@

We note that commutativity as assumed by Cyclone is an \emph{API level
  property}. There is no assumption as to how application cores synchronize to
access the shared key-value state or the underlying data structures for holding
key-value pairs. We emphasize that this means replicas ultimately have the same
content but not necessarily the same memory state.

@@

Our solution instead is to use a technique we term ganged operations. It is
based on the observation that in the failure free case, executing a batched
write across multiple control plane cores and KV store cores does not look very
much different from a synchronous shared memory barrier. The challenge is to
effectively manage failures in replication without causing participating cores
to stop making progress. The first step is to reconfigure RAFTs leader election
protocol to ensure that the leaders for all the physical logs are co-located on
the same host. Clients always dispatch batched writes to a fixed control plane
core (called the injection core) which is then responsible for forwarding the
request to the participating control plane cores (for corresponding physical
logs). The injection core uses DPDK's packet cloning primitives to avoid
dataplane work while forwarding copies of the same batched write packet to
multiple control plane cores. The injection core also adds a ``nonce'' - a
unique timestamp to the packet. In addition, the injection core adds a unique
view number to the packet containing the term numbers of all the participating
RAFT instances (read from shared memory of the co-located leaders).  The flow
for ganged operations is described in the pseudocode of
Figure~\ref{fig:ganged_ops}. We defer the discussion of how we generate the
nonce to later in this section. We also assume that a unique barrier is
allocated in shared memory for each ganged operation. We discuss later how this
is done without using dynamic allocation.

In the absence of failures Figure~\ref{fig:ganged_ops} is straightforward. Once
replication is complete on all participating physical logs, the corresponding KV
store cores execute \\{\tt event\_replicated\_ganged\_op}. They use a bitmask
for the barrier setting the appropriate bit for themselves. Once all KV store
cores have arrived at the barrier the operation is executed. A distinguished
leader core (we use the minumum numbered participating KV store core) is
responsible for sending back the response to the client. Two types of failure
need to be considered. First, replication on the physical log corresponding to
one of the non-leader cores could fail. The remaning cores would never progress
past the barrier. To solve this the leader core monitors the remaning cores for
their current view - which is the RAFT term of the last executed log entry. If
the view has moved past the view in the packet, replication failed on that
physical log. The leader then sets the bit for the core that will no longer
participate in the barrier to aid progress (a technique we borrowed from
lock-free data structures~\cite{lock-free}). It also sets a failed flag to
signal failure to the other participating cores. The second failure case is that
the leader core might not receive the operation if it fails during replication
from corresponding physical log. This case is detected in a similar manner by
the remaining cores by monitoring the published view from the leader core in
shared memory. We add that on a leadership change due to timeout we enqueue and
execute a nop on the physical log and all associated KV store cores to ensure
previous entries are committed - a corner case in the RAFT protocol.

@@

We now turn our attention from the network component of Cyclone to the storage
one by adding the second level flashlog. We evaluate the impact of adding
block storage to our hitherto pure packet switching scenario in
Figure~\ref{fig:flashlog}. Batching entries from the top level NVDIMM log to the
second level flashlog is clearly beneficial as adding flash storage at the
second level has almost no impact on peak performance in Cyclone for small
entries. The situation however changes for larger
entries. Figure~\ref{fig:flashlog_512} shows that using a 512 byte payload has a
significant impact on peak throughput - it drops to approximately 350K
ops/sec. This corresponds to around 50K 4KB IOPS to the SSD to write out the
flashlog pages. The peak for the drive is 160K IOPS using a queue depth
(concurrency) of 128. With 32 application threads we expect a lower peak
throughput around 40K IOPS explaining our bottleneck at 50K IOPS.  It is
possible to tune our observed performance further by aligning the flush boundary
to increase the number of outstanding requests - we do not do so in this paper,
keeping Cyclone agnostic to the exact size of entry being replicated. A final
point about Figure~\ref{fig:flashlog_512} is that once we are past the storage
bottleneck the latency spike is dramatic and large enough to trigger Cyclone's
failure detector and repeated retries from the clients. There are - therefore -
no points on the ``knee'' of the curve as in the pure packet switched one-level
log case.

@@

The final dimension we evaluate is of ganged operations. The primary purpose of
ganged operations is to avoid the need for distributed transactions to
manipulate what is a single shared memory image and therefore we were most
concerned about unloaded latency given the complexity of synchronizing different
replication quorums as well executing our rendezvous protocol on multiple
cores. We therefore setup an experiment where a single client - reflecting the
unloaded case - made ganged requests to the replicas. We varied the number of
cores participating in the request from one to the full complement of 32
application cores. Figure~\ref{fig:ganged} shows the results both using a single
level log as well as a two level log. The primary takeaway is that unloaded
latency increases slowly as we increase the number of active cores - to around
40 us from the baseline of 20 us. There are two causes of this. First, there is
a quicker rate of increase from 1 to 8 active cores, the reason being that this
corresponds to an increase in the number of physical logs that must be
synchronized on. There is some amount of synchronization involved in the
userspace DPDK driver for access to common resources for accessing the NIC and
this affects the ability to simultaneously issue replication messages on all
quorums. The smaller contribution to increasing latency that continues past the
8 core case is due to cacheline pingponging when executing the rendevous of
Figure~\ref{fig:ganged_ops}. Both these sources of latency drift could be
corrected using replication quorums mapped to dedicated NICs and more scalable
rendezvous designs (such as with machine aware broadcast
trees~\cite{broadcast_tree}). However we deemed the complexity of such
optimizations unnecessary. The added latency for most cases is well under the
extra round trip delay (including replication) for all phases of a 2PC.

@@

Cyclone takes a software only approach to improving replication performance
using commodity network hardware. At the other end of the spectrum, Consensus in
a box~\cite{consensus_box} implements the Zookeeper atomic broadcast protocol
together with a NIC on an FPGA. That work ignored durability focusing purely on
replication performance. Latency numbers are impressive due to the FPGA that cuts
out the path from CPU caches to the NIC and are reported to be as low as 3us for
a round trip from leader to follower replica compared to the 7us we observe with
DPDK. Notably however they reported 7us when using TCP/IP from the FPGA rather
than their own connection oriented network protocol. On the other hand peak
replication throughput appears to be 6M replications/sec on 66 Gbps of aggregate
network bandwith, comparable to the mark obtained by Cyclone on an aggregate 40
Gbps of network bandwidth. We believe the general applicability afforded by our
software only packet switching solution makes it a compelling alternative to an
FPGA based solution even given the larger latency.

@@

Cyclone provides a client library that automatically times out and retries
requests to deal with failure. Protection against repeated execution of the same
request is currently left to the key value store and client, since we felt that
exactly once semantics is optional when updates such as put operations are
idempotent. We have experimented with exactly once semantics as a basis for
distributed transactions across shards running Cyclone integrated key value
stores, along the lines of similar work~\cite{raft_lin}, using a fixed client
space and maintaining information about last seen sequence numbers from clients
in NVM. Distributed transactions on Cyclone shards however are outside the scope
of this paper.

Finally, we note that scaling a log horizontally for concurrency is also an idea
adopted by filesystems. NOVA~\cite{nova} is an example that uses a per-inode log
to improve concurrency when the filesystem is placed on NVM attached to the
memory bus. Cyclone also uses a distinct NVM log per control plane core in a
similar vein but does not maintain a namespace since the number of NVM logs is a
small fixed number equal to the number of control plane cores.

\section{Conclusion}
Persistent key value stores struggle today to provide high availability using a
write ahead log without a significant impact to performance. Cyclone solves this
problem by leveraging a small amount of directly attached non-volatile memory
that transforms the replication problem into a software packet switching
problem. Cyclone and its Rocksdb bindings are slated for release to the open
source community and we hope both the code and the ideas in the system itself
will help key value store builders and researchers overcome their logging
problem.
\newcommand\myurl[2]{\url{#1}}
\bibliographystyle{plain}
\bibliography{paper}
